---
title: "Reproducible Research with Verdifax"
author: "Your Lab"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: readable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE
)
```

# Why attest a computational result?

This R Markdown document shows how to take a small statistical analysis
from *"it runs on my machine"* to **cryptographically attested,
byte-identical on replay, bound to a declared environment fingerprint**.

The Verdifax pipeline does three things that go beyond what you can
do alone:

1. **Canonical seal.** Your result is normalized (RFC 8785 JCS) and
   hashed. Anyone with the same input, code, and declared environment
   can independently recompute the hash and check it matches.
2. **Reproducibility-context binding.** The orchestrator binds your
   declared runtime fingerprint (R version, pinned packages, git SHA,
   declared random seeds, platform, optional container hash) into the
   audit bundle as **Category 6**. The seal covers your declaration —
   not a guess.
3. **Transparency-log anchoring.** On Rekor-anchored deployments,
   every attestation publishes an immutable entry to the Sigstore
   transparency log. Erasing or rewriting the record after the fact
   is detectable.

## What you need

- A Verdifax API key — set `VERDIFAX_API_KEY` in your shell or
  `.Renviron`.
- The R SDK: `remotes::install_github("Verdifax/verdifax-sdk-r")`.

```{r install, eval = FALSE}
remotes::install_github("Verdifax/verdifax-sdk-r")
```

# 1. Set up a deterministic analysis

We fit a logistic regression on a small synthetic dataset with a
fixed seed. The model is intentionally trivial — the point is the
attestation pattern, not the statistics.

```{r model}
library(jsonlite)

set.seed(42)

n <- 200
x1 <- rnorm(n)
x2 <- rnorm(n)
x3 <- rnorm(n)
logit_p <- -0.4 + 0.9 * x1 - 0.6 * x2 + 0.3 * x3
y <- rbinom(n, 1, plogis(logit_p))

fit <- glm(y ~ x1 + x2 + x3, family = binomial)

result <- list(
  coefficients = round(as.numeric(coef(fit)), 8),
  coefficient_names = names(coef(fit)),
  null_deviance = round(fit$null.deviance, 8),
  residual_deviance = round(fit$deviance, 8),
  n_obs = as.integer(n)
)

# Canonical JSON form for hashing — sorted keys, no whitespace.
payload <- toJSON(result, auto_unbox = TRUE, digits = 8)
cat(payload, "\n")
```

# 2. Capture the research environment

`verdifax_capture_environment()` auto-detects the R runtime version,
every package version returned by `installed.packages()`, the git
commit SHA of the working directory (best-effort), the container image
hash (best-effort on Linux), and a GOOS/GOARCH-style platform
descriptor. You declare random seeds explicitly so the orchestrator
records exactly which seeds you claimed.

Auto-detection failures silently leave the field `NULL`. The
orchestrator records `NULL` as *"not declared"* rather than fabricating
a claim.

```{r capture}
library(verdifax)

ctx <- verdifax_capture_environment(declared_seeds = list(R = 42))
cat("runtime:       ", ctx$runtime_name, ctx$runtime_version, "\n")
cat("platform:      ", ctx$platform, "\n")
cat("git_commit_sha:", ctx$git_commit_sha %||% "(not detected)", "\n")
cat("# pinned deps: ", length(ctx$pinned_dependencies), "\n")
cat("declared seeds:", paste(ctx$random_seeds, collapse = ", "), "\n")
cat("declared flag: ", ctx$declared, "\n")
```

```{r helper-op, include = FALSE}
# Null-coalescing operator (used above for the cat() of git_commit_sha).
`%||%` <- function(x, y) if (is.null(x)) y else x
```

# 3. Attest the result

`verdifax_attest()` POSTs the payload to `/execute` and binds the
reproducibility context into the Category-6 section of the audit
bundle. The returned manifest carries the canonical `ManifestHash` —
the cryptographic seal of this analysis.

```{r attest, eval = nzchar(Sys.getenv("VERDIFAX_API_KEY"))}
client <- verdifax_client()  # reads VERDIFAX_API_URL / VERDIFAX_API_KEY

attestation <- verdifax_attest(
  client = client,
  payload = payload,
  program_id = strrep("0", 64),       # registry-authorized program ID
  route_id = "paper-figure-3",         # human-meaningful route label
  registry_record_hash = strrep("0", 64),  # §0 record hash; demo placeholder
  reproducibility_context = ctx
)

mh <- attestation$manifest$ManifestHash %||% attestation$manifest$manifest_hash
cat("run_id:       ", attestation$run_id, "\n")
cat("ManifestHash: ", mh, "\n")
```

# 4. Prove the pipeline is deterministic

`verdifax_verify_determinism()` runs the same payload through the
pipeline twice and reports whether both invocations produced
byte-identical canonical manifest hashes. The `deterministic` flag is
grounded on manifest-hash equality — the seal of the pipeline output.
Bundle-hash differences (when surfaced) reflect server-observed timing
variation and are labeled informational.

```{r verify, eval = nzchar(Sys.getenv("VERDIFAX_API_KEY"))}
det <- verdifax_verify_determinism(
  client = client,
  payload = payload,
  program_id = strrep("0", 64),
  route_id = "paper-figure-3-verify",
  registry_record_hash = strrep("0", 64),
  reproducibility_context = ctx
)

cat("deterministic:          ", det$deterministic, "\n")
cat("first.manifest_hash:    ", det$first$manifest_hash, "\n")
cat("second.manifest_hash:   ", det$second$manifest_hash, "\n")
cat("manifest_hash_differs:  ", det$diff$manifest_hash_differs, "\n")
cat("differing_fields (info):", paste(det$diff$differing_fields, collapse = ", "), "\n")
stopifnot(isTRUE(det$deterministic))
```

# 5. Re-knit this document

If you re-knit this document *without changing the input data, code,
or declared environment*, the `ManifestHash` printed in section 3 will
be **byte-identical** to the previous knit. That's the core
reproducibility guarantee Verdifax delivers: the same scientific
result always seals to the same cryptographic hash, and the seal is
bound to your declared environment fingerprint.

If anything *did* change — a package was bumped, a seed was different,
the input data was perturbed — the hash will differ, and any reader,
reviewer, or replicator will know.

# What to do with the manifest hash

- **Cite it in the paper.** Treat the manifest hash like a DOI for the
  computational result — short, unambiguous, machine-checkable.
- **Pin it in supplementary materials.** Reviewers can independently
  recompute it if they have the same inputs and declared environment.
- **Bind it to the artifact.** Sign your dataset or paper PDF over the
  manifest hash so the link is non-repudiable.

For audit-grade deployments, ask your administrator whether the
instance is **Rekor-anchored** — that adds an immutable Sigstore
transparency-log entry to every attestation, detectable to anyone
auditing the log.

# Session info

The chunk below records, alongside this knit, the full session info as
it would appear in a typical reproducibility appendix. Verdifax already
captures the same information cryptographically in section 2, but
including the human-readable form here is good hygiene.

```{r sessioninfo}
sessionInfo()
```
